Constraints on Cosmic-Ray Particle Populations, Magnetic Fields, and Dark Matter from VHE Gamma-Ray Observations of the Coma Cluster of Galaxies with VERITAS

### 2011-07-25 ms.tex
Some minor changes to text

Depending on the CR energy density, quantified by the fraction of energy in CRs to the thermal energy in a cluster, $\eta=\expval{ E_{\mathrm{CR}}/E_{\mathrm{th}}}$, an observable flux of high-energy gamma rays can be expected. This fraction could plausibly be as large as 50\% \citep{article:Ryu_etal:2003}, but a much smaller fraction of only a few percent is required in order to produce a gamma-ray flux observable with current generation gamma-ray telescopes.

@@

A third mechanism for gamma-ray production in a galaxy cluster is self-annihilation of a dark matter particle, e.g. a weakly interacting massive particle (WIMP).

@@

The analysis of this work follows the same procedure as described in detail in \citet{article:Abdo_etal:2009}. To eliminate albedo gamma rays from the Earth limb, a zenith angle cut of 105$^{\circ}$ is applied. The \emph{Fermi} Science Tools version v9r23p1 and Instrumental Response Functions (IRFs) P6\_V11\_Diffuse (a model of the spatial distribution of photon events calibrated  in flight) were used throughout this work. Only photons above 1 GeV are considered, where the constraints on the flux are most meaningful. The 99\% confidence level flux upper limits in the energy range 1-100 GeV, assuming both a unresolved, point-like or spatially extended emission are tabulated in Table \ref{table:fermi}. The gamma-ray emission is assumed to follow a power-law distribution with a photon index $\alpha=2.1$ and $2.5$.

\section{Gamma Ray Emission from Cosmic Rays}
We decided to adopt a multifaceted approach to constrain the CR-to-thermal pressure distribution in Coma with these upper limits. That includes (1) an analytic model derived from cosmological hydrodynamical simulations of the formation of galaxy clusters, (2) a simplified multi-frequency analytical approach that assumes a constant CR-to-thermal energy density and a power-law spectrum in momentum, and (3) an approach that uses the observed luminosity of the giant radio halo in Coma to place a lower limit on the expected $\gamma$-ray flux in the hadronic model -- independent of the magnetic field distribution. This translates into a minimum CR pressure which, if challenged by tight gamma-ray limits/detections, enables scrutinizing the hadronic interaction model of the formation of giant radio halos. Modelling the CR distribution through different techniques enable to bracket our ignorance about the underlying plasma
physics that shape the CR distribution and hence reflect the Bayesian priors that are imposed on the modelling \citep[see][for a discussion]{article:PinzkePfrommerBergstrom}.

@@

We adopt the electron density profile for the Coma cluster that has been inferred from ROSAT X-ray observations \citep{article:BrielHenryBohringer:1992} and model the spatial and spectral distribution of CRs according to the model by \citet{article:PinzkePfrommer:2010} that neglects the contribution of cluster galaxies. Figure ?? shows the expected spectrum of Coma, normalized to the total flux within the virial radius. [Add {\em Fermi} results]

@@

Table \ref{table:constraints} shows the expected gamma-ray fluxes within the regions of interest. For the most constraining regions tested (within an aperture of radius 0.2$^{\circ}$), those are a factor of 7.2 below the inferred upper limits (see Table \ref{table:results} and assuming a spectral index of $\alpha=2.3$ which matches the simulated one at energies $E=200$ GeV). Hence we scale the averaged relative CR pressure, $X_{\CR}$
% = \bra P_\CR \ket /\bra P_\rmn{th} \ket$ 
of our model by that factor, keep the spatial behavior and obtain an integrated limit of the relative CR pressure of $X_{\CR}<0.091$ within 0.2$^{\circ}$ that translates to a limit within the cluster virial radius of $X_\CR<0.13$ (solid lines of Figure \ref{fig:xcr}).

>>>>

\paragraph{rewrite this}
Depending on the CR energy density, quantified by the fraction of energy in CRs to the thermal energy in a cluster, $\eta=\expval{ E_{\mathrm{CR}}/E_{\mathrm{th}}}$, an observable flux of high-energy gamma rays can be expected. This fraction could plausibly be as large as 50\% \citep{article:Ryu_etal:2003}, but a much smaller fraction of only a few percent is required in order to produce a gamma-ray flux observable with current generation gamma-ray telescopes.

@@

A third mechanism for gamma-ray production in a galaxy cluster is self-annihilation of a dark matter particle, e.g. a weakly interacting massive particle (WIMP). This is particularly compelling because of the large amount of dark matter in a galaxy cluster.

@@

The analysis of this work follows the same procedure as described in detail in \citet{article:Abdo_etal:2009}. To eliminate albedo gamma rays from the Earth limb, a zenith angle cut of 105$^{\circ}$ is applied. The \emph{Fermi} Science Tools version v9r23p1 and Instrumental Response Functions (IRFs) P6\_V11\_Diffuse (a model of the spatial distribution of photon events calibrated  in flight) were used throughout this work. Only photons above 1 GeV are considered, where the constraints on the flux are most meaningful. The 95\% confidence level flux upper limits in the energy range 1-100 GeV, assuming both a unresolved, point-like or spatially extended emission are tabulated in Table \ref{table:fermi}. The gamma-ray emission is assumed to follow a power-law distribution with a photon index $\alpha=2.1$ and $2.5$.

\section{Gamma Ray Emission from Cosmic Rays}
We decided to adopt a multi-faceted approach to constrain the CR-to-thermal pressure distribution in the Coma cluster with the VERITAS and \emph{Fermi}-LAT upper limits from this work. This approach includes (1) an analytic model derived from cosmological hydrodynamical simulations of the formation of galaxy clusters, (2) a simplified multi-frequency analytical approach that assumes a constant CR-to-thermal energy density and a power-law spectrum in momentum, and (3) an approach that uses the observed luminosity of the giant radio halo in Coma to place a lower limit on the expected $\gamma$-ray flux in the hadronic model -- independent of the magnetic field distribution. This translates into a minimum CR pressure which, if challenged by tight gamma-ray limits/detections, enables scrutinizing the hadronic interaction model of the formation of giant radio halos. Modelling the CR distribution through different techniques enable to bracket our ignorance about the underlying plasma
physics that shape the CR distribution and hence reflect the Bayesian priors that are imposed on the modelling \citep[see][for a discussion]{article:PinzkePfrommerBergstrom}.

@@

We adopt the electron density profile for the Coma cluster that has been inferred from ROSAT X-ray observations \citep{article:BrielHenryBohringer:1992} and model the spatial and spectral distribution of CRs according to the model by \citet{article:PinzkePfrommer:2010} that neglects the contribution of cluster galaxies. Figure \ref{fig:spectrum} shows the expected spectrum of Coma, normalized to the total flux within the virial radius. [Add {\em Fermi} results]

@@

Table \ref{table:constraints} shows the expected gamma-ray fluxes within the regions of interest. For the most constraining regions tested (within an aperture of radius 0.2$^{\circ}$), those are a factor of 7.2 below the inferred upper limits (see Table \ref{table:results} and assuming a spectral index of $\alpha=2.3$ which matches the simulated one at energies $E=200$ GeV). Hence we scale the averaged relative CR pressure %$X_{\CR}$
% = \bra P_\CR \ket /\bra P_\rmn{th} \ket$ 
of our model by that factor, keep the spatial behavior and obtain an integrated limit of the relative CR pressure of $X_{\CR}<0.091$ within 0.2$^{\circ}$ that translates to a limit within the cluster virial radius of $X_\CR<0.13$ (solid lines of Figure \ref{fig:xcr}).
